Refunctionalization of Abstract Abstract Machines (Functional Pearl)

### 2018-07-04 paper/paper.tex
fix typos

\subsubsection{Allocating Addresses}
Up to this point, we have described neither allocating of addresses in stores, nor handling
of time stamps @Time@. In abstract interpretation, however, these are key ingredients
to realize analyses with different sensitivities, as well as to perform a finite
state space analysis \cite{Gilray:2016:ACP:2951913.2951936}.
To effectively approximate the run-time behavior, we introduce program \emph{contours} @time@ 
which are finite history of funciton calls till up to the current state \cite{Shivers:1991:SSC:115865.115884}.
The function @tick@ is used to refresh the ``time'' and get a ''new time''. 
We use a finite list of expressions (which are drawn from the control component) to encode the 
calling context history, and as we will see in Section~\ref{kcfainst}, by applying different 
@tick@ functions on the timestamp, we are able to obtain a family of analyses.

@@

The continuation and meta-continuation are added as arguments to @aeval@.
Before jumping into the details of refunctionalizing the stack structure and constructing
@new_cont@ and @new_mcont@, we have two important observations in the disentangled AAM
where the continuations were represented by data types:
1) When constructing the @new_ndk@, we observe that it explicitly uses @new_frames@ as one
argument of its top @NDCont@ object. The higer-order version of these continuations follows
the same pattern.
2) In the meta-continuation dispatching function @ndcontinue@, the size of the closures set is
decreasing every time while the other parts of meta-continuation are kept. This means in 
the higher-order version, we also need an ``ability'' to update the top frame of meta-continuation
and keep the rest as the same.

@@

But as mentioned above, defunctionalization and refunctionalization are more general concepts, 
with a wide range of uses. For example, \citet{Danvy:2001:DW:773184.773202} showed that 
defunctionalization is dual to Church encoding, and further case studies where known
higher-order and first-order implementations can be related through defunctionalization
and refunctionalization include backtracking algorithms (e.g., regular expression matcher), 
Dyck word recognizers, Dijkstra's shunting-yard algorithm, and even Quicksort 
\cite{Danvy:2006:RW:2171265.2171268, DANVY2009534, Danvy:2001:DW:773184.773202}. \\

@@

\textit{Abstracting Definitional Interpreters.}
Reynolds' seminal paper \emph{Definitional interpreters for higher-order programming languages} 
\cite{Reynolds:72, Reynolds:HOSC98-revisited} showed that in definitional interpreters, the
defined language inherits properties such as the order of evaluation from the defining
language, unless these properties are made explicit in the 
interpreter. With this insight,
\citet{darais2017abstracting} constructed abstract definitional interpreters which
automatically inherit the pushdown control-flow property from its defining language because
the defined language simply uses the call-stack model of the meta-language.
\citeauthor{darais2017abstracting}'s abstract definitional interpreters work on direct-style
$\lambda$-calculus, and are themselves written in monadic style. One of the advantages of monadic
abstract interpreters is modularity. Hence, deploying different sensitivities
or features can be achieved by just applying a different monad. Prior to that, \citet{Sergey:2013:MAI:2491956.2491979}
presented a monadic abstract interpreter for small-step semantics.

>>>>

\subsubsection{Allocating Addresses}
Up to this point, we have described neither allocating of addresses in stores, nor handling
of time stamps @Time@. In abstract interpretation, however, these are key ingredients
to realize analyses with different sensitivities, as well as to perform a finite
state space analysis \cite{Gilray:2016:ACP:2951913.2951936}.
To effectively approximate the run-time behavior, we introduce program \emph{contours} @time@ 
which are finite history of function calls till up to the current state \cite{Shivers:1991:SSC:115865.115884}.
The function @tick@ is used to refresh the ``time'' and get a ''new time''. 
We use a finite list of expressions (which are drawn from the control component) to encode the 
calling context history, and as we will see in Section~\ref{kcfainst}, by applying different 
@tick@ functions on the timestamp, we are able to obtain a family of analyses.

@@

The continuation and meta-continuation are added as arguments to @aeval@.
Before jumping into the details of refunctionalizing the stack structure and constructing
@new_cont@ and @new_mcont@, we have two important observations in the disentangled AAM
where the continuations were represented by data types:
1) When constructing the @new_ndk@, we observe that it explicitly uses @new_frames@ as one
argument of its top @NDCont@ object. The higher-order version of these continuations follows
the same pattern.
2) In the meta-continuation dispatching function @ndcontinue@, the size of the closures set is
decreasing every time while the other parts of meta-continuation are kept. This means in 
the higher-order version, we also need an ``ability'' to update the top frame of meta-continuation
and keep the rest as the same.

@@

But as mentioned above, defunctionalization and refunctionalization are more general concepts, 
with a wide range of uses. For example, \citet{Danvy:2001:DW:773184.773202} showed that 
defunctionalization is dual to Church encoding, and further case studies where known
higher-order and first-order implementations can be related through defunctionalization
and refunctionalization include backtracking algorithms (e.g., regular expression matcher), 
Dyck word recognizer, Dijkstra's shunting-yard algorithm, and even Quicksort 
\cite{Danvy:2006:RW:2171265.2171268, DANVY2009534, Danvy:2001:DW:773184.773202}. \\

@@

\textit{Abstracting Definitional Interpreters.}
Reynolds' seminal paper \emph{Definitional interpreters for higher-order programming languages} 
\cite{Reynolds:72, Reynolds:HOSC98-revisited} showed that in definitional interpreters, the
defined language inherits properties such as the order of evaluation from the defining
language, unless these properties are made explicit in the 
interpreter. With this insight,
\citet{darais2017abstracting} constructed abstract definitional interpreters which
automatically inherit the pushdown control-flow property from its defining language because
the defined language simply uses the call-stack model of the meta-language.
\citeauthor{darais2017abstracting}'s abstract definitional interpreters work on direct-style
$\lambda$-calculus, and are themselves written in monadic style. One of the advantages of monadic
abstract interpreters is modularity. Hence, deploying different sensitivities
or features can be achieved by just applying different monads. Prior to that, \citet{Sergey:2013:MAI:2491956.2491979}
presented a monadic abstract interpreter for small-step semantics.
